---
title: "Final Project"
author: "Blue"
date: "2023-08-10"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

NEW DATA SET

```{r}
library(readr)
#read in fire data
fire_data <- read_csv('mapdataall.csv')

View(mapdataall.csv)
```

```{r}
library(tigris)
library(ggplot2)

school <- school_districts(state = "CA")
counties <- counties(state = "CA")
counties$sum_fire

#testing out to make sure tigris is working


ggplot()+
  geom_sf(data = counties)

```



```{r}
#creating a bunch of different geographical plots
library(readr)
library(dplyr)
library(tigris)
library(ggplot2)
library(sf)
library(magrittr)

# Load the county boundaries for California
counties <- counties(state = "CA")

# Calculate the areas of all counties
county_areas <- st_area(counties)

# Convert the areas to acres
acres <- county_areas / 4046.86

# Create a new dataframe with county names and their corresponding areas
county_areas_df <- data.frame(NAME = counties$NAME, Acres = acres)


counties <- counties(state = "CA")
fire_data <- read_csv("mapdataall.csv")

fire_data_sf <- fire_data %>%
  st_as_sf(coords = c("incident_longitude", "incident_latitude")) %>%
  st_set_crs(4326) %>%
  filter(incident_acres_burned >= 5000)

ggplot() +
  geom_sf(data = counties, fill = NA, color = "black") +
  geom_sf(data = fire_data_sf, aes(color = incident_acres_burned), size = 2) +
  scale_color_viridis_c(option = "magma", direction = -1) +
  theme_minimal() +
  labs(color = "Acres Burned") +
  coord_sf(xlim = c(-125, -114), ylim = c(32, 43))  # california

fire_data %<>%
  st_as_sf(coords = c("incident_longitude", "incident_latitude")) 

fire_data %<>%
  st_set_crs(st_crs(counties))

fire_data %<>%
  filter(`incident_acres_burned` >= 5000)


counties %<>%
  st_join(fire_data)

counties %<>%
  group_by(GEOID)
counties %<>%
  summarise(sum_fire = sum(incident_acres_burned, na.rm = T))

ggplot(counties)+
  geom_sf(aes(fill = sum_fire))+
  scale_fill_viridis_c(option = "magma", direction = -1)+
  theme_minimal()+
  labs(fill = "sum of acres burned")

```

```{r}
library(readr)
library(dplyr)
library(tigris)
library(ggplot2)
library(sf)
library(ggplot2)

# Load the county boundaries for California

counties <- counties(state = "CA")

# Calculate the areas of all counties

county_areas <- st_area(counties)
acres <- county_areas / 4046.86
county_areas_df <- data.frame(NAME = counties$NAME, Acres = acres)
county_areas_df <- county_areas_df %>%
  mutate(
    Acres = as.numeric(gsub("\\D", "", Acres))
  )
fire_data <- read_csv("mapdataall.csv")

fire_data_sf <- fire_data %>%
  st_as_sf(coords = c("incident_longitude", "incident_latitude")) %>%
  st_set_crs(4326) %>%
  filter(incident_acres_burned >= 5000)#only fires that burned 5000+acres

ggplot() +
  geom_sf(data = counties, fill = NA, color = "black") +
  geom_sf(data = fire_data_sf, aes(color = incident_acres_burned), size = 2) +
 
  theme_minimal() +
  labs(color = "Acres Burned") +
  coord_sf(xlim = c(-125, -114), ylim = c(32, 43))  # california image

fire_data %<>%
  st_as_sf(coords = c("incident_longitude", "incident_latitude"))  #converts fire_data into a spatial data frame

fire_data %<>%
  st_set_crs(st_crs(counties)) #This sets the CRS of the fire_data to match the CRS of the counties spatial data.

counties %<>%
  st_join(fire_data) #matching the fire_data fires to their specific counties

counties %<>%
  left_join(county_areas_df, by = c("NAME" = "NAME"))

counties %<>%
  group_by(GEOID) %>%
  summarise(sum_fire_ratio = sum(incident_acres_burned, na.rm = TRUE) / sum(Acres, na.rm = TRUE)) #calculating the proportion of acres burned to total acres in each county.

#ggplot of the acres burned/acres in each county. With a linear scale.
ggplot(counties) +
  geom_sf(aes(fill = sum_fire_ratio)) +

  theme_minimal() +
  labs(fill = "Sum of Acres Burned / Total County Acres")  # Set the legend label


ggplot(counties) +
  geom_sf(aes(fill = sum_fire_ratio)) +
  scale_fill_viridis_c(trans = "log10") +  # Apply a logarithmic color scale since it was not understandable with a linear scale (too polar)
  theme_minimal() +
  labs(fill = "Sum of Acres Burned / Total County Acres")  # Set the legend label

  
  

```


conext for the outlier: the Thomas fire in Ventura county was massive. It burned about 282,000 acres and took 6 months to put out compltetly. It was started by power lines and it as able to spread so effectively because of strong winds and because there was a lot of thick, dry brush in the area that was perfect wildfire fuel.

backstory on some of the counties, chapter 6 scatter plot acres vs wter quality, change over time, compare counties to future counties.