---
title: "Homework 3 394"
author: "Blue"
date: "2025-02-02"
output: html_document
---

a. (Shelley) Below are the last five lines of Shelley’s poem Ozymandias

"My name is Ozymandias, king of kings:
Look on my works, ye Mighty, and despair!"
Nothing beside remains. Round the decay
Of that colossal wreck, boundless and bare
The lone and level sands stretch far away

Let X denote the length of a word which is randomly selected from those lines. What is X(boundless)?

x(boundless) = 9 since there are 9 letters in "boundless"

b. (Free throws) As part of the warm up drill, each player on a basketball team is required to shoot free
throws until a basket is made. If Amy has a 80% success rate at the foul line, write the PMF of the
random variable X that describes the number of misses before she completes the drill. Assume that
individual throws constitute independent events.
Write the PMF of X as a formula first, and then write a brief explanation below.


$$f(x) = (1-P(S))^x*P(S)^1$$
$$f(x) = .2^x*.8$$
this formula account for x misses before a single shot with a probability of 80% is made. each event is independent of the last. .2 is raised to the power of x since there is a 20% chance of a miss x times and .8 is raised to the first power since only one shot will be scored.

2)(Random walk) Suppose a particle moves 4 steps along the x-axis, starting at 0. At each step, it moves
randomly one unit to the right or to the left.

A)If you copy and paste the following question into ChatGPT: How many elements are in the sample space?
Explain your answer very briefly , you will get the answer shown in the file HW3problem-chatgpt.pdf
which is located in the COURSEMATERIALS/HOMEWORK folder.
The number of elements is correctly calculated, however, there is an error made in the explanation for
why. In your own words, describe the error and how you would fix it.

ChatGPT claims that the number of elements in the sample space is 2^4 = 16 since each step has two possible outcomes and you are doing it 4 times. An error in it's explanation is that it says because each step is independent, you can calculate the sample space with 2^4. Independence does not matter when computing the sample space and chatGPT should not have mentioned it.

b. Let E denote the event that the particle ends up at x = 0. Write the outcomes in E and calculate
P (E). Since this is a binomial distribution since there are a fixed # of trials, each event is independent, probabilities stay constant, and only 2 outcomes per trial. This example fits all the critera so we can use the formula $$\binom{4}{2} \times (0.5)^2 \times (0.5)^2$$ to get .375. .375*16 = 6 so we know that there are 6 outcomes were x = 0. These possibilities are (RRLL), (RLRL), (RLLR), (LLRR), (LRLR), (LRRL)

C. Now suppose the random variable X denotes the position of the particle after 4 steps. Write its PMF
in a tabular format. I have created a partial table for you to fill in. Each row should contain a possible
value x, then list the outcomes from the original sample space that give that possible value and the
probability. (Hint: in part b you calculated P (X = 0).
$$\begin{array}{|c|c|c|}
\hline
\text{X} & \text{outcomes from S} & f(X) \\
\hline
-4 & 1 & 0.0625 \\
-2 & 4 & 0.25   \\
0  & 6 & 0.375  \\
2  & 4 & 0.25   \\
4  & 1 & 0.0625 \\
\hline
\end{array}
$$



d) Create a code chunk and write code to draw the probability histogram of X, the position of the particle
after 4 steps. Please see the slidedeck on chapter 6 for sample code. Be sure to give labels and a title.
Do not use any code that you have not been taught in this class

```{r}
library(tidyverse)


bud <- tibble(
  x = c(-4,-2,0,2,4),
  f = c(.0625, .25,.375,.25,.0625)
)

# make probability histogram
ggplot(data = bud,
mapping = aes(x=x, y=f) ) +
geom_col(width = 1, fill = "gold", color = "purple", alpha = 0.5)+
labs(x = "ending x", y = "probability", title = "equal probability histogram")





```
```{r}


(1/3)**4 #prob of -4
choose(4,3)*((1/3)**3)*((2/3)**1) #prob of -2
choose(4,2)*((1/3)**2)*((2/3)**2) # prob of 0
choose(4,1)*((1/3)**1)*((2/3)**3) #prob of 2
(2/3)**4 # prob of 4




```

$$\begin{array}{|c|c|c|}
\hline
\text{X} & \text{outcomes from S} & f(X) \\
\hline
-4 & 1 & 0.01234568 \\
-2 & 4 & 0.09876543   \\
0  & 6 & 0.2962963  \\
2  & 4 & 0.3950617   \\
4  & 1 & 0.1975309 \\
\hline
\end{array}
$$
Since the probability of the outcome being right increased, the probability of ending up with a larger number also increase. The most likely outcome is now ending up at x = 2 after 4 trials. we are still assuming that each event is independent and the probabilites stay constant at 1/3 for left and 2/3 for right. This is still a binomial distribution which makes calculating the probabilites with the above formulas in the code chunk possible.


```{r}
library(tidyverse)


bud <- tibble(
  x = c(-4,-2,0,2,4),
  f = c(0.01234568, 0.09876543,0.2962963,0.3950617,0.1975309)
)

# make probability histogram
ggplot(data = bud,
mapping = aes(x=x, y=f) ) +
geom_col(width = 1, fill = "gold", color = "purple", alpha = 0.5)+
labs(x = "ending x", y = "probability", title = "unequal probability histogram")

```


4. (Discrete Uniform) Suppose X is a discrete uniform random variable on the integers 1, 2, . . . , n, that is,
it has PMF
fX (x) = P (X = x) = k, x = 1, 2, . . . , n
where k is some unknown number.

a. Find k.


$$\sum_{x=1}^{n} P(X = x) = 1 = \sum_{x=1}^{n} k$$


$$kn = 1$$
$$k = 1/n$$


b. Define the random variable Y = −X Write the PMF of Y along with a brief explanation.

fY (y) = P (Y = y) = k = 1/n, y = -1, -2, . . . , -n

since Y = -X all the values of Y will just become the negative version of x


