---
title: "Week5socAss"
author: "Blue"
date: "2023-07-31"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1) Run a topic model with 4 topics on the “guardian_co_river.csv” dataset.

```{r}
library(readr)
library(dplyr)
library(tidytext)
 library(topicmodels)
co_river <- read.csv("guardian_co_river.csv")

tidy_co <- co_river %>%
  unnest_tokens(word, body_text) %>%
  anti_join(stop_words)
  
  tidy_co_counts <- tidy_co %>%
  count(id, word, sort = TRUE)
  
  co_dtm <- tidy_co_counts %>%
  cast_dtm(id, word, n)
  
# lda 
co_lda <- LDA(co_dtm, 4, control = list(seed = 1))

```
```{r}
# colorado topics
library(readr)
library(dplyr)
library(tidytext)
 library(topicmodels)
#install.packages('reshape2')
library(reshape2)
co_topics <- tidy(co_lda, matrix = "beta")

co_top_terms <- co_topics %>%
  group_by(topic) %>%
  slice_max(beta, n = 10) %>% 
  ungroup() %>%
  arrange(topic, -beta) %>%
  mutate(term = reorder_within(term, beta, topic))
```
Plot the top words for each topic, similar to what is done in section 5.2. What would you label these topics?

```{r}
library(ggplot2)
# plot our topics!
ggplot(co_top_terms, aes(beta, term, fill = factor(topic))) +
geom_col(show.legend = FALSE) +
facet_wrap(~ topic, scales = "free") +
scale_y_reordered()
```

1) nature
2) national park
3) urban/city
4) politics

Try running the LDA again, this time with a different number of topics. Does the new topic model make more or less sense to you?
```{r}
library(readr)
library(dplyr)
library(tidytext)
 library(topicmodels)
#install.packages('reshape2')
library(reshape2)
co_lda2 <- LDA(co_dtm, 6, control = list(seed = 1))

co_topics2 <- tidy(co_lda2, matrix = "beta")

co_top_terms2 <- co_topics2 %>%
  group_by(topic) %>%
  slice_max(beta, n = 10) %>% 
  ungroup() %>%
  arrange(topic, -beta) %>%
  mutate(term = reorder_within(term, beta, topic))

library(ggplot2)
# plot our topics!
ggplot(co_top_terms2, aes(beta, term, fill = factor(topic))) +
geom_col(show.legend = FALSE) +
facet_wrap(~ topic, scales = "free") +
scale_y_reordered()


```
This topic model makes less sense to me since there are many repeat words across different topics such as "canyon" and 2 extremely similar political topics.


Using whichever topic model made the most sense to you, create a dataset similar to co_documents in section 5.3. Choose two articles to examine in more detail. Discuss the topic probabilities.

```{r}
library(readr)
library(dplyr)
library(tidytext)
 library(topicmodels)
#install.packages('reshape2')
library(reshape2)
co_documents_og <- tidy(co_lda, matrix = "gamma")
head(co_documents_og)
#environment/2010/sep/06/ship-fresh-water-alaska-india
#us-news/2021/apr/05/arizona-water-one-percenters
co_documents2 <- co_documents_og %>%
  filter(gamma > 0.95) %>%
  filter(document == "environment/2010/sep/06/ship-fresh-water-alaska-india") 
   
co_documents2

co_documents <- co_documents_og %>%
  filter(gamma > 0.95) %>%
  filter(document == "us-news/2021/apr/05/arizona-water-one-percenters") 
   


```
Both of these articles were filtered so that they were at least 95% focussed on one topic. They will both be practically only talking about one tepic in the article. I wasn't sure if you wanted me to make a topic model for these two articles or not, if you did I'm happy to re do this section.

Considering what we know about the Colorado River and water rights in the U.S. (from readings and class), are we surprised by the topics generated by these news articles? Describe how your results compared to expectations.

The topics generated focussed on nature, national parks, urban/city, and politics.
These topcs make complete sense since the colorado river was shrinking which brings in the natural topic, and there are politics surrounding the shrinkage(such as mexico not getting their alloted water and how different states are colliding), urban cities will be impacted since they have to reduce water usage, The national park section does confuse me a bit but when looking closer to the other words in the topic its more of a second nature section.