---
title: "Lab 04: Data Visualization and Mapping"
subtitle: "SOC 225: Data & Society"
author: "Blue Knutson"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---

# Agenda
- Getting to know the data set
- Visualize covariations using ggplot2
  - scatterplots
  - scales
  - colors
  - facets
- Make maps
  - points and popups
  - colors and legends

# 1. Getting to know `gapminder`

For the first half of the lab, we'll use the `gapminder` data set and `ggplot2` package to learn the basics about data visualization. `ggplot2` is a core package under `tidyverse`. The data frame `gapminder` will become available once you have loaded the package.

For the second half of the lab, we'll use the Seattle Airbnb listings from Inside Airbnb.

**Question 1.1. Install the `gapminder` package in the console. Load `gapminder` and `tidyverse` in the below code chunk.** 
```{r}
# install.packages("gapminder")
library(gapminder)
data(gapminder)
```

Before we start, let's take a look at the `gapminder` data.
**Question 1.2. Use str() to take a look at data frame `gapminder`.**

```{r}
str(gapminder)
```
**Question 1.3. From checking the structure of the data frame, how many different types of data do you find? Which are continuous, and which are categorical?**
*Hint at the bottom*
* 2 types: continuous and categorical
* 2 categorical:country, continent
* 4 continuous: year, lifeExp, pop, gdpPercap



**Question 1.4. How many observations does the data frame contain?**

* 1704

You may also notice that `gapminder` has a nested/hierarchical structure: `year` in `country` in `continent`. These are panel data!


**Question 1.5. Now, create a subset on country Algeria.**

```{r}
library(tidyverse)
algeria <- gapminder %>% filter(country=="Algeria")
View(algeria)

```

Read about the origins of the `gapminder` data set by typing `?gapminder`, and by looking at the source of the data: https://www.gapminder.org/data/.

# 2. Visualize covariations using ggplot2

A great way of plotting is to use the `ggplot2` package. The core idea underlying this package is the *layered grammar of graphics*: we can break up elements of a plot into pieces and combine them. 

`ggplot2` graph objects consist of two primary components:

- Layers, the components of a graph. 
  - We *add* layers to a `ggplot2` object using `+`.
  - This includes lines, shapes, and text.
- Aesthetics, which determines how the layers appear. 
  - We *set* aesthetics using *arguments* (e.g. color="red") insider layer functions.
  - This includes locations, colors, and sizes.

We'll use `ggplot2` to learn how to show the *covariation* between variables. Different from bar charts or histograms where only *variation* of a single variable is displayed, in this lab we focus on how different variables may vary together. 

## Scatterplots

Scatterplots are a good way to visualize the relationship between two variables, and to look for outliers.

```{r}
theme_set(theme_minimal()) # setting minimal theme as defalt

p <- # use the assignment operator to save a plot!
  ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) + # use the gapminder data, set x to the gdp and y to the life expectancy
  geom_point() # plot points (x,y) for each country in the data set

p # by calling the object 'p' we can display the graph
```

**Question 2.1. Make a scatterplot comparing the population and the life expectancy, assign it to an object named 'q'**

```{r}
q<-ggplot(gapminder, aes(x=pop,y=lifeExp))+geom_point()
  

q

```

## Scales

It appears that countries with higher per capita gdp may have higher life expectancy. Data points are however lumped together within the gdpPercap 0-30,000 range. To better see our data, we can transform the x-axis into a log scale.

```{r}
p + # taking our old graph, the '+' lets us add to it
  scale_x_log10() # add our scale
```

**Question 2.2. add a log scale to your graph q**

```{r}
q %>% + scale_x_log10()
```

**Question 2.3. what do you think accounts for the distinctive shape of your scatterplot?**
The lines are there because the life expectancy is constantly going up/related. If they were seperated by country the points would line up pretty will
## Colors

In addition to the x and y axes, another aesthetic available to use is `color`. This means we can look at how up to three variables change together!

For example, we can color data points by continent, a categorical variable: 

```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) + # same as p above but adding color
  geom_point() + # plot the points
  scale_x_log10() # add the same scale
```

If you map a continuous/numeric variable onto color, `ggplot2` will pick a gradient scale: 

```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) + 
  geom_point() + 
  scale_x_log10()
```

**Question 2.4. Remake plot q by choosing a continuous variable to add colors**

```{r}
ggplot(gapminder, aes(x = pop, y = lifeExp, color = gdpPercap)) + 
  geom_point() + 
  scale_x_log10()

```

**Question 2.5. Remake plot q by choosing a categorical variable to add colors that explain the interesting shapes**

```{r}
q<-
  ggplot(gapminder, aes(x=pop, y=lifeExp, color=continent))+geom_point()+scale_x_log10()
q

```


## Facets

Let's take a different way of breaking down these data by continent. This time, we'll *facet* the data into "small multiple" plots.

To do this, we add a new layer with `facet_wrap`. Note: the syntax is slightly different! You use a tilde (`~`) before the variable name.

(Why? Because you can facet by more than one variable. In R, this syntax is called a *formula*.)

```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) + 
  geom_point() + 
  scale_x_log10() + 
  facet_wrap(~continent) 
```

**Question 2.6: Add a continent facet wrap to q**

```{r}
q<-
  ggplot(gapminder, aes(x=pop,y=lifeExp, color =continent))+geom_point()+facet_wrap(~continent)
q
```

# 3 Exercise: Life expectancy over time

Instead of looking at the relationship between life expectancy and GDP, now we'll look at changes in life expectancy over time. You can use the last code block for all the questions.

## 3.1: Line plot of life expectancy by year 

**Create a plot where `x = year` and `y = lifeExp`.** This time, use a new geom: `geom_line()` instead of `geom_point()`. Initially, it won't look quite right.  


## 3.2: The `group` aesthetic 

**You need to tell the plot to group the lines by country.** To do this, you'll need a new aesthetic, `group = country`. Incorporate this into your plot. Does it look more reasonable now?

## 3.3: Facet and interpret

**Finally, facet by continent.** Does life expectancy seem to have increased over time everywhere? Do you see any dips or decreases? 

```{r}
ggplot(gapminder, aes(x = year, y = lifeExp, group = country)) + 
  geom_line() + 
  scale_x_log10() + 
  facet_wrap(~continent) 

```

# 4. Maps

You can create maps in R using latitude and longitude data. This uses a new package called `leaflet` (https://rstudio.github.io/leaflet/), which you'll need to install in the console. 

The Inside Airbnb data has latitudes and longitudes for each listing, so we'll use that.

**Question 4.1: Install leaflet, load it, and then read the file 'data/listings.csv' and name it airbnb_data**

```{r}
#install.packages("leaflet")
airbnb_data <- read.csv("listings.csv")
```


## Points and popups

`leaflet` relies on the pipe (`%>%`) to add layers to maps. 

```{r}
library(leaflet)
leaflet(airbnb_data) %>% # begin by passing the data to leaflet
  addTiles() %>% # add the map files to the plot, 
                 # leaflet automatically uses the latitude and longitude data 
                 # to find the right map
  addCircles(popup = ~name) # add circles for each listing, 
                            # by adding the popup argument, we can click on a 
                            # cirle to show the name
```

**Question 4.2: Make a leaflet plot that only includes only listings with a price over $200 and shows the price when the circle is clicked**

```{r}
library(leaflet)
library(tidyverse)


leaflet(filter(airbnb_data, price>200)) %>% # begin by passing the data to leaflet
  addTiles() %>% # add the map files to the plot, 
                 # leaflet automatically uses the latitude and longitude data 
                 # to find the right map
  addCircles(popup = ~name) # add circles for each listing, 
                            # by adding the popup argument, we can click on a 
                            # cirle to show the name


```

## Colors and legends: qualitative/categorical

Colors are a bit less automatic than with `ggplot2`. You need to create a palette using one of the provided functions, and then use that palette for your data and the legend. 

```{r}
# We're going to start by making a smaller data frame to use for our visualization
example_data <- 
  airbnb_data %>% # assign the small dataframe the name 'example_data'
  filter(neighbourhood == "University District") # only include listings from the UDistrict
 
# this palette is based on the type of room
# "Set1" is a qualitative palette name
room_type_pal <- colorFactor("Set1", example_data$room_type) 

leaflet(example_data) %>% # make a leaflet plot with our example data
  addTiles() %>% # add the map tiles
  addCircles(popup = ~name, # add circles which we can click to see the names
             color = ~room_type_pal(room_type)) %>% # color the circles with our palette
  addLegend(pal = room_type_pal, values = ~room_type) # add a legend so we know what's what
```


**Question 4.3: Use the same code from above, but map a neighborhood group of your choice**

```{r}
# We're going to start by making a smaller data frame to use for our visualization
example_data <- 
  airbnb_data %>% # assign the small dataframe the name 'example_data'
  filter(neighbourhood == "Green Lake") # only include listings from the UDistrict
 
# this palette is based on the type of room
# "Set1" is a qualitative palette name
room_type_pal <- colorFactor("Set1", example_data$room_type) 

leaflet(example_data) %>% # make a leaflet plot with our example data
  addTiles() %>% # add the map tiles
  addCircles(popup = ~name, # add circles which we can click to see the names
             color = ~room_type_pal(room_type)) %>% # color the circles with our palette
  addLegend(pal = room_type_pal, values = ~room_type) # add a legend so we know what's what
```


## Colors and legends: numeric/continuous

You should use different color palettes for categorical vs numeric data. You've got a couple options for plotting numeric data:

- `colorNumeric`: linear mapping of numbers onto a color gradient. For example, `colorNumeric("RdPu", data$variable)`.

- `colorBin`: bins by values, so each color spans the same numeric range. For example, `colorBin("RdPu", data$variable, bins = 5)`.  

- `colorQuantile`: bins by quantiles, so each color has the same number of data points. For example, `colorQuantile("RdPu", data$variable, n = 5)`.

The examples above use a red-purple gradient ("RdPu") as the color palette. Type `RColorBrewer::brewer.pal.info` into your console for a full list of possible palettes.

```{r}
example_data <- 
  airbnb_data %>% 
  filter(neighbourhood == "University District") 

price_pal <- colorNumeric("RdPu", example_data$price)

leaflet(example_data) %>%
  addTiles() %>%
  addCircles(popup = ~name, 
             color = ~price_pal(price)) %>%
  addLegend(pal = price_pal, values = ~price)
```

**Question 4.4: Use the same code from above, but make a palette based on quantiles instead.**

```{r}
example_data <- 
  airbnb_data %>% 
  filter(neighbourhood == "University District") 

price_pal <- colorQuantile("RdPu", example_data$price, n = 5)

leaflet(example_data) %>%
  addTiles() %>%
  addCircles(popup = ~name, 
             color = ~price_pal(price)) %>%
  addLegend(pal = price_pal, values = ~price)
```

**Question 4.5: Compare the two maps of prices in the UDistrict. Which mapping of prices to colors---linear mapping by value or binning by quantile---do you think is more useful here, and why? There isn't a right answer.**
Linear mapping is more useful because when people budget for an airbnb they want to know the price of a home, not just that its relatively expensive or cheap relative to other airbnbs.


*Hint 1.3*
Variables `country` and `continent` are **factor** variables. Factor variables are categorical data with an underlying numerical representation. 

# References
Kieran Healy, [Data Visualization: A Practical Introduction](http://socviz.co/)

Charles Lanfear, [Introduction to R for Social Scientists](https://clanfear.github.io/CSSS508/)


