---
title: "LAB 9 Submission sheet"
output:
  html_document:
    df_print: paged
---


Modify this file only to add your answers. Please, do not delete any code from this file.

```{r load-packages, echo=FALSE, message=FALSE, results = "hide"}
if(!require("tidyverse")) install.packages("tidyverse")
if(!require("openintro")) install.packages("openintro")
if(!require("statsr")) install.packages("statsr")
if(!require("broom")) install.packages("broom")
library(tidyverse)
library(openintro)
library(statsr)
library(broom)
library(infer)

us_adults <- tibble(
  climate_change_affects = c(rep("Yes", 62000), rep("No", 38000))
)

set.seed(1)
n <- 60
samp <- us_adults %>%
  sample_n(size = n)

yrbss <- yrbss %>% 
  mutate(physical_3plus = if_else(physically_active_7d > 2, "yes", "no"))  %>%
  na.exclude()

obs_diff <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

null_dist <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

t_obs <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  calculate(stat = "t", order = c("yes", "no"))

t_dist <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  assume(distribution = "t")
```

### Question 1

<!---- Answer to Question 1 starts here ----->

Write here your answer to question 1. Write the code, if any, in the chunk, and comments/interpretations below the chunk.

```{r question1}
# Insert Code Here

samp %>%
  specify(response = climate_change_affects, success = "Yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.90)

```



<!---- Answer to Question 1 ends here ----->


### Question 2

<!---- Answer to Question 2 starts here ----->

Write here your answer to question 2. Write the code, if any, in the chunk, and comments/interpretations below the chunk.

```{r question2}
# Insert Code Here



```

the CI using the central limit theorom is (0.2140537 0.4526129) and using the bootstrap it's (0.55	0.7833333) so the boot strap was higher than the central limit theorom.
<!---- Answer to Question 2 ends here ----->


### Question 3

<!---- Answer to Question 3 starts here ----->

Write here your answer to question 3. Write the code, if any, in the chunk, and comments/interpretations below the chunk.

```{r question3}
# Insert Code Here



```

H(0): mu(no) - mu(yes) = 0

H(A): mu(no)- mu(yes) != 0
<!---- Answer to Question 3 ends here ----->




### Question 4

<!---- Answer to Question 4 starts here ----->

We observe a value more extreme than the observed value with a probability of 2*p for a two tailed distribution using the null distribution .  This relates to the p value because that probability is 2 times the p value. p value = values more extreme than observed/ value in null. In this case there are no values more extreme than the P value so we don't observe any greater than the P value.

Write here your answer to question 4. Write the code, if any, in the chunk, and comments/interpretations below the chunk.

```{r question4}
# Insert Code Here



```


<!---- Answer to Question 4 ends here ----->



### Question 5

<!---- Answer to Question 5 starts here ----->

Write here your answer to question 5. Write the code, if any, in the chunk, and comments/interpretations below the chunk.

```{r question5}
# Insert Code Here



```

The permutation model computes the p value off of the raw data while the theoretical model converts to the t distribution to compute the p value. I prefer the theoretical model beacause it allows us to get a better estimate of the p value since the permutation method gave us a p value of 0 which is just false under a normal distrbution while the theoretical model was able to give a more realisitc answer.
<!---- Answer to Question 5 ends here ----->





